Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-25 paper.tex
minor edits to the conclusion

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
  \label{thm:SD-bisim}
  Assume $C$ implements Lazy D Cast ADT
  and $C$ is a monoid
  and $cast(T,l,T)=id(T)$,
  $s_{1} \in \text{\effCEK{C}}$,
  $s_{2} \in \text{\ineffCEKD}$,
  $s_{1} \approx s_{2}$,
\begin{enumerate}
  \item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
%  \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
  \item If $\judgeSreduce{C}{s_1}{s_3}$ 
    then $\judgeDreduce{s_2}{s_4}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_4$
  \item If $\judgeDreduce{s_2}{s_4}$
    then $\judgeSreduce{C}{s_1}{s_3}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_3$
\end{enumerate}
{\color{red}
  Jeremy's attempt to reformulate in a more standard way that
  remains equivalent:
\begin{enumerate}
\item If $\judgeSreduce{C}{s_1}{s_3}$ then
  $\judgeDreducePlus{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$.
\item If $\judgeDreduce{s_2}{s_4}$ then
  $\judgeSreduce{C}{s_1}{s_3}$,
  $\judgeDreduceTrans{s_2}{s_5}$, and
   $s_3 \approx s_5$ for some $s_3$ and $s_5$.
\item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
\end{enumerate}
}
Or diagrammatically, 
  \[\begin{array}{clclc}
  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
  \rotatebox[origin=c]{90}{$\approx$} 
  & & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
  s_{2} & 
  \longrightarrow_{\text{\ineffCEKD}} & s_2' &
  \longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
  \end{array}\]
\end{lemma}
\begin{proof}
  Part (1) follow from the definition of $s \approx s$.
  Part (2) and (3) are more interesting. 
  By the definition of $s \approx s$, it is straightforward to prove 
  $\judgeSreduce{C}{s_1}{s_3}$ if and only if $\judgeDreduce{s_2}{s_4}$.
  So we can prove two parts together by showing that 
  \[
  \text{if}\; s_{1} \approx s_{2}
  \;\text{and}\;\judgeSreduce{C}{s_1}{s_3}
  \;\text{and}\;\judgeDreduce{s_2}{s_4}
  \;\text{then}\; s_3 \approx^{*} s_4.
  \]
  To put it another way, our goal is to proving 
  one reduction step in \effCEK{C} corresponds to
  one or more steps in \ineffCEKD.  
  Although there are many cases, the general idea is the same.
  Consider
  \[
  \sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{
    wrap(\kOOcar{\kappa})
  }
  \approx
  \sOOreturn{v_2}{\kOOcar{\kappa}}
  \]

@@

\begin{lemma}[Strong Bisimulation among $\mathcal{S}(\cdot)$]
  \label{thm:CEKS-bisim}
  Assume $C_1$ and $C_2$ implements Lazy D Cast ADT
  $s_{1} \in \text{\effCEK{C_1}}$,
  $s_{2} \in \text{\effCEK{C_2}}$,
  $s_{1} \approx s_{2}$,
  \begin{enumerate}
    \item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
    \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
    \item If $\judgeSreduce{C_1}{s_1}{s_3}$ 
    then $\judgeSreduce{C_2}{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$
    \item If $\judgeSreduce{C_2}{s_2}{s_4}$ 
    then $\judgeSreduce{C_1}{s_1}{s_3}$ and $s_3 \approx s_4$ for some $s_3$
  \end{enumerate}
% 
  Or diagrammatically, 
  \[\begin{array}{clc}
  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_3\\
  \rotatebox[origin=c]{90}{$\approx$} 
  & & \rotatebox[origin=c]{90}{$\approx$} \\
  s_{2} & 
  \longrightarrow_{\text{\ineffCEKD}} & s_4 \\
  \end{array}\]
\end{lemma}
\begin{proof} This proof is straightforward. The key ideas are undoing 
sequencing with the property (2) of Lazy D Cast ADT, and handling all possibly 
uses of $cast(T,l,T)$ with property (3)-(11).
\end{proof}

@@

In this paper we presented a new cast representation, hypercoercions, in two 
flavors, Lazy D and Lazy UD. Hypercoercions have a structurally recursive 
composition and have a more compact memory representation in comparison to 
coercions in normal form.

We also present steps toward the first framework for proving the correctness of 
cast representations.
%
The framework includes abstract data types for cast representations (Cast ADT 
and Lazy D Cast ADT), an parameterized abstract machine \effCEK{C}, and a 
theorem states that \effCEK{C} is equivalent to $D$ if $C$ is a Lazy D Cast 
ADT.
%
Lazy D Cast ADT refines Cast ADT. We conjecture that Lazy D threesomes, and 
Lazy D coercions in normal form are also Lazy D Cast ADT.

Finally, we proved that Lazy D hypercoercions ($H$) is a Lazy D Cast ADT. By 
using our framework, we shown that it is a correct cast representation because 
$S(H)$ is equivalent to $D$.

>>>>

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
  \label{thm:SD-bisim}
  Assume $C$ implements Lazy D Cast ADT
  and $C$ is a monoid
  and $cast(T,l,T)=id(T)$,
  $s_{1} \in \text{\effCEK{C}}$,
  $s_{2} \in \text{\ineffCEKD}$,
  $s_{1} \approx s_{2}$,
\begin{enumerate}
  \item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
  \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
  \item If $\judgeSreduce{C}{s_1}{s_3}$ 
    then $\judgeDreduce{s_2}{s_4}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_4$
  \item If $\judgeDreduce{s_2}{s_4}$
    then $\judgeSreduce{C}{s_1}{s_3}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_3$
\end{enumerate}
{\color{red}
  Jeremy's attempt to reformulate in a more standard way that
  remains equivalent:
\begin{enumerate}
\item If $\judgeSreduce{C}{s_1}{s_3}$ then
  $\judgeDreducePlus{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$.
\item If $\judgeDreduce{s_2}{s_4}$ then
  $\judgeSreduce{C}{s_1}{s_3}$,
  $\judgeDreduceTrans{s_4}{s_5}$, and
   $s_3 \approx s_5$ for some $s_3$ and $s_5$.
\item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
\end{enumerate}
}
Or diagrammatically, 
  \[\begin{array}{clclc}
  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
  \rotatebox[origin=c]{90}{$\approx$} 
  & & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
  s_{2} & 
  \longrightarrow_{\text{\ineffCEKD}} & s_2' &
  \longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
  \end{array}\]
\end{lemma}
\begin{proof}
  Part (1) and (2) follow from the definition of $s \approx s$.
  Part (3) and (4) are more interesting. 
  By the definition of $s \approx s$, it is straightforward to prove 
  $\judgeSreduce{C}{s_1}{s_3}$ if and only if $\judgeDreduce{s_2}{s_4}$.
  So we can prove two parts together by showing that 
  \[
  \text{if}\; s_{1} \approx s_{2}
  \;\text{and}\;\judgeSreduce{C}{s_1}{s_3}
  \;\text{and}\;\judgeDreduce{s_2}{s_4}
  \;\text{then}\; s_3 \approx^{*} s_4.
  \]
  To put it another way, our goal is to proving 
  one reduction step in \effCEK{C} corresponds to
  one or more steps in \ineffCEKD.  
  Although there are many cases, the general idea is the same.
  Consider
  \[
  \sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{
    wrap(\kOOcar{\kappa})
  }
  \approx
  \sOOreturn{v_2}{\kOOcar{\kappa}}
  \]

@@

\begin{lemma}[Strong Bisimulation among $\mathcal{S}(\cdot)$]
  \label{thm:CEKS-bisim}
  Assume $C_1$ and $C_2$ implements Lazy D Cast ADT
  $s_{1} \in \text{\effCEK{C_1}}$,
  $s_{2} \in \text{\effCEK{C_2}}$,
  $s_{1} \approx s_{2}$,
  \begin{enumerate}
    \item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
    \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
    \item If $\judgeSreduce{C_1}{s_1}{s_3}$ 
    then $\judgeSreduce{C_2}{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$
    \item If $\judgeSreduce{C_2}{s_2}{s_4}$ 
    then $\judgeSreduce{C_1}{s_1}{s_3}$ and $s_3 \approx s_4$ for some $s_3$
  \end{enumerate}
% 
  Or diagrammatically, 
  \[\begin{array}{clc}
  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_3\\
  \rotatebox[origin=c]{90}{$\approx$} 
  & & \rotatebox[origin=c]{90}{$\approx$} \\
  s_{2} & 
  \longrightarrow_{\text{\ineffCEKD}} & s_4 \\
  \end{array}\]
\end{lemma}
\begin{proof} This proof is straightforward. The key ideas are undoing 
  uses of $seq$ with the property (2) of Lazy D Cast ADT, and handling all
  uses of $cast$ with properties (3)-(11).
\end{proof}

@@

In this paper we presented a new cast representation, hypercoercions,
in two flavors, Lazy D and Lazy UD. Hypercoercions have a structurally
recursive composition operator and have a more compact memory
representation in comparison to coercions in normal form.

We also present steps toward the first framework for proving the
correctness of cast representations.
%
The framework includes abstract data types for cast representations
(Cast ADT and Lazy D Cast ADT), a parameterized abstract machine
\effCEK{C}, and a theorem which states that \effCEK{C} is equivalent
to $D$ when $C$ is a Lazy D Cast ADT.
%
The Lazy D Cast ADT refines the Cast ADT. We conjecture that Lazy D
threesomes, and Lazy D coercions in normal form are instances of the
Lazy D Cast ADT.

Finally, we proved that Lazy D hypercoercions ($H$) is a Lazy D Cast
ADT. By using our framework, we showed that it is a correct cast
representation because $S(H)$ is equivalent to $D$.
